Triptych

### 2019-10-25 paper/main.tex
Incomplete special-soundness proof

We sometimes use index subscript notation of the form $i_j$ to indicate the $j$ digit of $i$, where such a decomposition of $i$ is taken base $n$ with padded length $m$.
That is, for $n=2$, this indexes a binary decomposition.
This notation will be specified explicitly where confusion may occur.

\subsection{Prover}
The prover algorithm accepts the following:
\begin{itemize}
\item a set of outputs $\{M_k\}_{k=0}^{N-1} \subset \G$
\item a set of corresponding amount commitments $\{P_k\}_{k=0}^{N-1} \subset \G$
\item a set of private keys $\{r^{(u)}\}_{u=0}^{w-1} \subset \F$
\item a set of commitment amounts and blinding factors $\{(s^{(u)},a_u)\}_{u=0}^{w-1} \subset \F$
\item a set $\{l^{(u)}\}_{u=0}^{w-1} \subset \F$ such that:
\begin{itemize}
\item each $M_{l^{(u)}} = r^{(u)}G$ has its private key known to the prover
\item each $P_{l^{(u)}} = s^{(u)}G + a_uH$ has its amount and blinding factor known to the prover
\end{itemize}
\item a set of new amount commitments $\{Q_j\}_{j=0}^{T-1} \subset \G$
\item a set of commitment amounts and blinding factors $\{(t_j,b_j)\}_{j=0}^{T-1} \subset \F$ such that:
\begin{itemize}
\item each $Q_j = t_jG + b_jH$ has its amount and blinding factor known to the prover
\item $\sum_{u=0}^{w-1} a_j - \sum_{j=0}^{T-1} b_j = 0$
\end{itemize}
\end{itemize}

The prover then performs these steps:
\begin{itemize}
\item Define $\left\{J^{(u)}\right\}_{u=0}^{w-1} \subset \G$ such that $J^{(u)} \equiv (r^{(u)})^{-1}\hp(M_{l_u})$.
\item Select random $r_A \in \F$ and $\left\{a^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1} \subset \F$.
Define $A \equiv \com(a,r_A)$.
\item Define $\left\{\sigma^{(u)}_{j,i}\right\}_{i,j,u=0}^{n-1,j-1,w-1} \subset \F$ such that $\sigma^{(u)}_{j,i} \equiv \delta\left(l^{(u)}_j,i\right)$ (using our decomposition notation), and choose random $r_B \in \F$.
Define $B \equiv \com(\sigma,r_B)$.
\item Select random $r_C \in \F$, and define $C \equiv \com(a(1-2\sigma), r_C)$.
\item Select random $r_D \in \F$, and define $D \equiv \com(-a^2, r_D)$.
\item Select random $\left\{\rho^{(u)}_j\right\}_{j,u=0}^{m-1,w-1}, \left\{\overline{\rho}^{(u)}_j\right\}_{j,u=0}^{m-1,w-1} \subset \F$.
\item For each $0 \leq u < w$, define coefficients $\left\{p^{(u)}_{k,j}\right\}_{k,j=0}^{N-1,m-1}$ such that $$p^{(u)}_k(x) \equiv \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} = \delta\left(l^{(u)},k\right)x^m + \sumj p^{(u)}_{k,j}x^j$$ (using our decomposition notation).
Then define $p_{k,j} \equiv \sumu p^{(u)}_{k,j}$ and $p_k(x) \equiv \sumu p^{(u)}_k(x)$ accordingly.
\item Define aggregation coefficients $\{\mu_k\}_{k=0}^{N-1}$ using a $k$-indexed $\hs$ operation, containing all public parameters and proof elements.
\item Define $\{X_j\}_{j=0}^{m-1} \subset \G$ such that: $$X_j \equiv \sumk p_{k,j}\mu_kM_k + \sumu \rho^{(u)}_jG$$
\item Define $\{Y_j\}_{j=0}^{m-1} \subset \G$ such that: $$Y_j \equiv \sumk p_{k,j}\mu_k\hp(M_k) + \sumu \rho^{(u)}_jJ^{(u)}$$
\item Define $\{Z_j\}_{j=0}^{m-1} \subset \G$ such that: $$Z_j \equiv \sumk p_{k,j}P_k + \sumu \overline{\rho}^{(u)}_jG$$
\item Define the Fiat-Shamir transcript challenge $x \in \F$ computed using all public parameters and these elements (with set notation dropped for brevity): $$M,P,Q,J,A,B,C,D,X,Y,Z$$
\item Define $\left\{f^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1}$ such that $f^{(u)}_{j,i} \equiv \sigma^{(u)}_{j,i}x + a^{(u)}_{j,i}$.
\item Define $z_A \equiv r_A + xr_B$ and $z_C \equiv xr_C + r_D$.
\item Define $\left\{z^{(u)}_R\right\}_{u=0}^{w-1} \subset \F$ such that: $$z^{(u)}_R \equiv \mu_{l^{(u)}}r^{(u)}x^m - \sumj \rho^{(u)}_jx^j$$
\item Define: $$z_S \equiv x^m\left( \sumu s^{(u)} - \sum_{j=0}^{T-1} t_j \right) - \sumj \left(x^j \sumu \overline{\rho}^{(u)}_j\right)$$
\end{itemize}

The proof consists of these elements (with set notation dropped for brevity):
$$J,A,B,C,D,X,Y,Z,f,z_A,z_C,z_R,z_S$$
A proof consists of $3m + w + 4$ elements of $\G$, and $w\left[ m(n-1) + 1 \right] + 3$ elements of $\F$.

\subsection{Verifier}
The verifier accepts the following:
\begin{itemize}
\item a set of outputs $\{M_k\}_{k=0}^{N-1}$
\item a set of corresponding amount commitments $\{P_k\}_{k=0}^{N-1}$
\item a proof generated by a prover
\end{itemize}

@@

@@

To show the correctness of the $X$ check:
\begin{eqnarray*}
&& \sumk \mu_kM_k \left[ \sumu \left( \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} \right) \right] - \sumj x^jX_j - \sumu z^{(u)}_RG \\
&=& \sumk \mu_kM_k p_k(x) - \sumj x^j \left( \sumk p_{k,j}\mu_kM_k + \sumu \rho^{(u)}_jG \right) - \sumu z^{(u)}_RG \\
&=& \sumk \mu_kM_k \left( p_k(x) - \sumj x^j p_{k,j} \right) - \sumj x^j \sumu \rho^{(u)}_jG - \sumu z^{(u)}_RG \\
&=& \sumk \mu_kM_k \left[ \sumu \delta\left( l^{(u)},k \right) \right] - \sumj x^j \sumu \rho^{(u)}_jG - \sumu\left[ \mu_{l^{(u)}}r^{(u)}x^m - \sumj \rho^{(u)}_jx^j \right]G \\
&=& x^m\sumu \mu_{l^{(u)}}r^{(u)}G - \sumj x^j \sumu \rho^{(u)}_jG - x^m\sumu \mu_{l^{(u)}}r^{(u)}G + \sumj x^j \sumu \rho^{(u)}_jG \\
&=& 0
\end{eqnarray*}

The $Y$-check proceeds similarly, but with more irritating algebra:
\begin{eqnarray*}
&& \sumk \mu_k\hp(M_k) \left[ \sumu \left( \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} \right) \right] - \sumj x^jY_j - \sumu z^{(u)}_RJ^{(u)} \\
&=& \sumk \mu_k\hp(M_k) p_k(x) - \sumj x^j \left( \sumk p_{k,j}\mu_k\hp(M_k) + \sumu \rho^{(u)}_jJ^{(u)} \right) - \sumu z^{(u)}_RJ^{(u)} \\
&=& \sumk \mu_k\hp(M_k) \left( p_k(x) - \sumj x^j p_{k,j} \right) - \sumj x^j \sumu \rho^{(u)}_jJ^{(u)} - \sumu z^{(u)}_RJ^{(u)} \\
&=& \sumk \mu_k\hp(M_k) \left[ \sumu \delta\left( l^{(u)},k \right) \right] - \sumj x^j \sumu \rho^{(u)}_j(r^{(u)})^{-1}\hp(M_{l^{(u)}}) - \sumu\left[ \mu_{l^{(u)}}r^{(u)}x^m - \sumj \rho^{(u)}_jx^j \right](r^{(u)})^{-1}\hp(M_{l^{(u)}}) \\
&=& x^m\sumu \mu_{l^{(u)}}\hp(M_{l^{(u)}}) - \sumj x^j \sumu \rho^{(u)}_j(r^{(u)})^{-1}\hp(M_{l^{(u)}}) - x^m\sumu \mu_{l^{(u)}}\hp(M_{l^{(u)}}) + \sumj x^j \sumu \rho^{(u)}_j(r^{(u)})^{-1}\hp(M_{l^{(u)}}) \\
&=& 0
\end{eqnarray*}

@@

>>>>

We sometimes use index subscript notation of the form $i_j$ to indicate the $j$ digit of $i$, where such a decomposition of $i$ is taken base $n$ with padded length $m$:
$$\sum_{j=0}^m i_j n^j = i$$
This notation will be specified explicitly where confusion may occur.

\subsection{Prover}
The prover algorithm accepts the following:
\begin{itemize}
\item a set of outputs $\{M_k\}_{k=0}^{N-1} \subset \G$
\item a set of corresponding amount commitments $\{P_k\}_{k=0}^{N-1} \subset \G$
\item a set of private keys $\{r^{(u)}\}_{u=0}^{w-1} \subset \F$
\item a set of commitment amounts and blinding factors $\{(s^{(u)},a_u)\}_{u=0}^{w-1} \subset \F$
\item a set of linking tags $\left\{J^{(u)}\right\}_{u=0}^{w-1} \subset \G$
\item a set $\{l^{(u)}\}_{u=0}^{w-1} \subset \F$
\begin{itemize}
\item each $M_{l^{(u)}} = r^{(u)}G$ has its private key known to the prover
\item each $P_{l^{(u)}} = s^{(u)}G + a_uH$ has its amount and blinding factor known to the prover
\item each $J^{(u)} = (r^{(u)})^{-1}\hp(M_{l_u})$
\end{itemize}
\item a set of new amount commitments $\{Q_j\}_{j=0}^{T-1} \subset \G$
\item a set of commitment amounts and blinding factors $\{(t_j,b_j)\}_{j=0}^{T-1} \subset \F$ such that:
\begin{itemize}
\item each $Q_j = t_jG + b_jH$ has its amount and blinding factor known to the prover
\item $\sum_{u=0}^{w-1} a_j - \sum_{j=0}^{T-1} b_j = 0$
\end{itemize}
\end{itemize}

The prover then performs these steps:
\begin{itemize}
\item Select random $r_A \in \F$ and $\left\{a^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1} \subset \F$.
Define $A \equiv \com(a,r_A)$.
\item Define $\left\{\sigma^{(u)}_{j,i}\right\}_{i,j,u=0}^{n-1,j-1,w-1} \subset \F$ such that $\sigma^{(u)}_{j,i} \equiv \delta\left(l^{(u)}_j,i\right)$ (using our decomposition notation), and choose random $r_B \in \F$.
Define $B \equiv \com(\sigma,r_B)$.
\item Select random $r_C \in \F$, and define $C \equiv \com(a(1-2\sigma), r_C)$.
\item Select random $r_D \in \F$, and define $D \equiv \com(-a^2, r_D)$.
\item Select random $\left\{\rho^{(u)}_j\right\}_{j,u=0}^{m-1,w-1}, \left\{\overline{\rho}^{(u)}_j\right\}_{j,u=0}^{m-1,w-1} \subset \F$.
\item For each $0 \leq u < w$, define coefficients $\left\{p^{(u)}_{k,j}\right\}_{k,j=0}^{N-1,m-1}$ such that $$p^{(u)}_k(x) \equiv \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} = \delta\left(l^{(u)},k\right)x^m + \sumj p^{(u)}_{k,j}x^j$$ (using our decomposition notation).
Then define $p_{k,j} \equiv \sumu p^{(u)}_{k,j}$ and $p_k(x) \equiv \sumu p^{(u)}_k(x)$ accordingly.
\item Define aggregation coefficients $\{\mu_k\}_{k=0}^{N-1}$ using a $k$-indexed $\hs$ operation, containing all public parameters and proof elements.
\item Define $\{X_j\}_{j=0}^{m-1} \subset \G$ such that: $$X_j \equiv \sumk p_{k,j}\mu_kM_k + \sumu \rho^{(u)}_jG$$
\item Define $\{Y_j\}_{j=0}^{m-1} \subset \G$ such that: $$Y_j \equiv \sumk p_{k,j}\mu_k\hp(M_k) + \sumu \rho^{(u)}_jJ^{(u)}$$
\item Define $\{Z_j\}_{j=0}^{m-1} \subset \G$ such that: $$Z_j \equiv \sumk p_{k,j}P_k + \sumu \overline{\rho}^{(u)}_jG$$
\item Define the Fiat-Shamir transcript challenge $x \in \F$ computed using all public parameters and these elements (with set notation dropped for brevity): $$M,P,Q,J,A,B,C,D,X,Y,Z$$
\item Define $\left\{f^{(u)}_{j,i}\right\}_{i=1,j,u=0}^{n-1,m-1,w-1}$ such that $f^{(u)}_{j,i} \equiv \sigma^{(u)}_{j,i}x + a^{(u)}_{j,i}$.
\item Define $z_A \equiv r_A + xr_B$ and $z_C \equiv xr_C + r_D$.
\item Define $\left\{z^{(u)}_R\right\}_{u=0}^{w-1} \subset \F$ such that: $$z^{(u)}_R \equiv \mu_{l^{(u)}}r^{(u)}x^m - \sumj \rho^{(u)}_jx^j$$
\item Define: $$z_S \equiv x^m\left( \sumu s^{(u)} - \sum_{j=0}^{T-1} t_j \right) - \sumj \left(x^j \sumu \overline{\rho}^{(u)}_j\right)$$
\end{itemize}

The proof consists of these elements (with set notation dropped for brevity):
$$A,B,C,D,X,Y,Z,f,z_A,z_C,z_R,z_S$$
A proof consists of $3m + w + 4$ elements of $\G$, and $w\left[ m(n-1) + 1 \right] + 3$ elements of $\F$.

\subsection{Verifier}
The verifier accepts the following:
\begin{itemize}
\item a set of outputs $\{M_k\}_{k=0}^{N-1}$
\item a set of corresponding amount commitments $\{P_k\}_{k=0}^{N-1}$
\item a set of linking tags $\{J^{(u)}\}_{u=0}^{w-1}$
\item a proof generated by a prover
\end{itemize}

@@

Proofs follow similarly to those in \cite{bootle}.

@@

To show the correctness of the $X$ check:
\begin{eqnarray*}
&& \sumk \mu_kM_k \left[ \sumu \left( \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} \right) \right] - \sumj x^jX_j - \sumu z^{(u)}_RG \\
&=& \sumk \mu_kM_k p_k(x) - \sumj x^j \left( \sumk p_{k,j}\mu_kM_k + \sumu \rho^{(u)}_jG \right) - \sumu z^{(u)}_RG \\
&=& \sumk \mu_kM_k \left( p_k(x) - \sumj x^j p_{k,j} \right) - \sumj x^j \sumu \rho^{(u)}_jG - \sumu z^{(u)}_RG \\
&=& \sumk \mu_kM_k \left[ x^m \sumu \delta\left( l^{(u)},k \right) \right] - \sumj x^j \sumu \rho^{(u)}_jG - \sumu\left[ \mu_{l^{(u)}}r^{(u)}x^m - \sumj \rho^{(u)}_jx^j \right]G \\
&=& x^m\sumu \mu_{l^{(u)}}r^{(u)}G - \sumj x^j \sumu \rho^{(u)}_jG - x^m\sumu \mu_{l^{(u)}}r^{(u)}G + \sumj x^j \sumu \rho^{(u)}_jG \\
&=& 0
\end{eqnarray*}

The $Y$-check proceeds similarly, but with more irritating algebra:
\begin{eqnarray*}
&& \sumk \mu_k\hp(M_k) \left[ \sumu \left( \prod_{j=0}^{m-1} f^{(u)}_{j,k_j} \right) \right] - \sumj x^jY_j - \sumu z^{(u)}_RJ^{(u)} \\
&=& \sumk \mu_k\hp(M_k) p_k(x) - \sumj x^j \left( \sumk p_{k,j}\mu_k\hp(M_k) + \sumu \rho^{(u)}_jJ^{(u)} \right) - \sumu z^{(u)}_RJ^{(u)} \\
&=& \sumk \mu_k\hp(M_k) \left( p_k(x) - \sumj x^j p_{k,j} \right) - \sumj x^j \sumu \rho^{(u)}_jJ^{(u)} - \sumu z^{(u)}_RJ^{(u)} \\
&=& \sumk \mu_k\hp(M_k) \left[ x^m \sumu \delta\left( l^{(u)},k \right) \right] - \sumj x^j \sumu \rho^{(u)}_j(r^{(u)})^{-1}\hp(M_{l^{(u)}}) - \sumu\left[ \mu_{l^{(u)}}r^{(u)}x^m - \sumj \rho^{(u)}_jx^j \right](r^{(u)})^{-1}\hp(M_{l^{(u)}}) \\
&=& x^m\sumu \mu_{l^{(u)}}\hp(M_{l^{(u)}}) - \sumj x^j \sumu \rho^{(u)}_j(r^{(u)})^{-1}\hp(M_{l^{(u)}}) - x^m\sumu \mu_{l^{(u)}}\hp(M_{l^{(u)}}) + \sumj x^j \sumu \rho^{(u)}_j(r^{(u)})^{-1}\hp(M_{l^{(u)}}) \\
&=& 0
\end{eqnarray*}

@@

\subsection{Special soundness}
We claim that the protocol is $(m+1)$-special sound, where $m > 1$.
To show this, we construct an extractor that, given $m+1$ valid responses to $m+1$ distinct verifier challenges for the same initial statement, produces a valid witness.
In particular, we produce a modified witness for the following relation that is based on the information presented in the prover algorithm:
\begin{multline*}
\left\{ \{M_k\}_{k=0}^{N-1},\{P_k\}_{k=0}^{N-1},\{J^{(u)}\}_{u=0}^{w-1},\{Q_j\}_{j=0}^{T-1} \subset \G \: ; \: \left( \{l^{(u)}\}_{u=0}^{w-1}, r, y \right) : r = \sum_{u=0}^{w-1} \hs(l^{(u)},M,P,Q)r^{(u)} \text{ and } \right. \\
\left. M_{l^{(u)}} = r_uG \; \forall \: u \in [0,w) \text{ and } r^{(u)}J^{(u)} = \hp(M_{l^{(u)}}) \; \forall \: u \in [0,w) \text{ and } \sum_{u=0}^{w-1} P_{l^{(u)}} - \sum_{j=0}^{T-1} Q_j = yG \right\}
\end{multline*}

Collision resistance of $\hs$ means that, with high probability, knowledge of $r$ is equivalent to knowledge of $\{r^{(u)}\}$.
The combination of the discrete logarithm problem on $\G$ and the collision-resistance of $\hp$ allows for double-spend enforcement using the linking tags $\{J^{(u)}\}$.
The binding property of Pedersen commitments ensures that, for properly constructed commitments, knowledge of $y$ implies that the sum of values of input commitments $\{P_{l^{(u)}}\}$ balances the sum of output commitments $\{Q_j\}$.
We may assume here that such commitments are correctly constructed, as this is shown by range proofs outside the scope of the proving system.

Suppose that for a given statement, we have a set of $(m+1)$ distinct verifier challenges $\{x_e\}_{e=0}^m$ corresponding to distinct valid responses of this form:
$$\left\{ \{{}_ef^{(u)}_{j,i}\}, \{{}_ez^{(u)}_R\}, {}_ez_S \right\}_{e=0}^m$$
From the $3$-special soundness in \cite{bootle} and $m > 1$ we have valid extractions $\{\sigma^{(u)}_{j,i}\}_{u=0}^{w-1}$ and $\{a^{(u)}_{j,i}\}_{u=0}^{w-1}$, and the Pedersen binding property ensures that (with high probability) we have:
$${}_ef^{(u)}_{j,i} = \sigma^{(u)}_{j,i}x_e + a^{(u)}_{j,i} \; \forall \: e \in [0,m]$$
Using the extracted values, compute
$$p_k(x) \equiv \sum_{u=0}^{w-1}\left[ \prod_{j=0}^{m-1} \left( \sigma^{(u)}_{j,k}x + a^{(u)}_{j,k} \right) \right]$$
for all $k \in [0,N)$.

We have seen that $p_k$ is of degree $m$ only when $k \in \{l^{(u)}\}_{u=0}^{w-1}$.
Hence there exist coefficients $\{\overline{X}_j,\overline{Y}_j,\overline{Z}_j\}_{j=0}^{m-1}$, computed uniquely from the statement and extracted values, such that the verification equations are of the following form:
\begin{eqnarray*}
x^m \sumu \mu_{l^{(u)}}M_{l^{(u)}} + \sumj x^j\overline{X}_j &=& \left( \sumu z^{(u)}_R \right)G \\
x^m \sumu \mu_{l^{(u)}}\hp(M_{l^{(u)}}) + \sumj x^j\overline{Y}_j &=& \sumu z^{(u)}_RJ^{(u)} \\
x^m \sumu P_{l^{(u)}} + \sumj x^j\overline{Z}_j - x^m \sum_{j=0}^{T-1} Q_j &=& z_SG
\end{eqnarray*}

Construct a Vandermonde matrix $V$ where the $e$ row is the vector $(1,x_e,\ldots,x^m_e)$.
Since all $x_e$ are distinct, the rows of $V$ span $\F^{m+1}$; hence there exist weights $\{\theta_e\}_{e=0}^m$ such that the resulting linear combination of rows produces the vector $(0,\ldots,0,1)$.
That is, $\sum_{e=0}^m \theta_ex^j = \delta(j,m)$.

For each verification equation, we can therefore build a linear combination over $e$. For the $X$-check:
$$\sumu \mu_{l^{(u)}} M_{l^{(u)}} = \sum_{e=0}^m \theta_ex_e^m \left( \sumu \mu_{l^{(u)}} M_{l^{(u)}} \right) + \sum_{e=0}^m \theta_e \left( \sumj x_e^j \overline{X}_j \right) = \left[ \sum_{e=0}^m \theta_e \left( \sumu {}_ez^{(u)}_R \right) \right]G$$
The rightmost bracketed double sum is therefore, with high probability, equal to $r = \mu_{l^{(u)}} r^{(u)}$.

The $Z$-check proceeds similarly:
$$\sumu P_{l^{(u)}} - \sum_{j=0}^{T-1} Q_j = \sum_{e=0}^m \theta_ex_e^m \left( \sumu P_{l^{(u)}} - \sum_{j=0}^{T-1} Q_j \right) + \sum_{e=0}^m \theta_e \left( \sumj x_e^j \overline{Z}_j \right) = \left( \sum_{e=0}^m \theta_e {}_ez_S  \right)G$$
Hence we have $y \equiv \sum_{e=0}^m \theta_e {}_ez_S$.

Unfortunately, the same technique does not appear to apply for the $Y$-check. Doing so obtains the following:
$$\sumu \mu_{l^{(u)}} \hp(M_{l^{(u)}}) = \sum_{e=0}^m \theta_ex_e^m \left( \sumu \mu_{l^{(u)}} \hp(M_{l^{(u)}}) \right) + \sum_{e=0}^m \theta_e \left( \sumj x_e^j \overline{Y}_j \right) = \sumu \left( \sum_{e=0}^m \theta_e {}_ez^{(u)}_R \right) J^{(u)}$$
It is unclear how to associate this with the required witness extraction.

\textbf{Therefore, the special-soundness proof is incomplete!}
